---
title: "ANOVA thesis MB quantitative data nematods in nodule crevices"
author: "el mario"
date: "22/04/2019"
output: html_document
---
# RQ: Do the Boxcore samples (BC's) differ amongst each other in their Nematod content (amount)?
#     Do the subsamples per BC differ more from each other than between BC's?
#     1-way ANOVA: Quantitative analysis


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install/load packages
```{r}
library("readxl")
library("car")
library("ggpubr")
library("ggplot2")
```

# Loading and inspecting the dataset

```{r}
# read dataset:

my_data <- read_excel(path = "/Users/lailabuydens/Desktop/excel_scripts_thesis/Nodule_crevice.xlsx", sheet = 1, col_names = TRUE, col_types = NULL, na = "", skip = 0)
head(my_data) # check dataset
```

```{r}
# Visualizing the data

ggboxplot(my_data, x = "location_ID", y = "square_root_transform", color = "location", palette = c("#FF8000", "#A5DF00", "#B40431"),
          order = c("B6SO2", "B4SO3", "B4NO1"),
          ylab = expression(bold("Total density")), xlab = expression(bold("Location")))

```



# Before ANOVA: check for assumptions (= data is normally distributed + variances are equal between locations)

```{r}

## testing for normality: visual test with QQ-plots and Shapiro-Wilk normality test PER LOCATION (H0: normal distribution => p<0.05: reject hypothesis => no normal distribution)

# per location check for normal distribution of raw data
B6S02 <- my_data$raw_data_total[1:15]
B4S03 <- my_data$raw_data_total[16:30]
B4N01 <- my_data$raw_data_total[31:45]

# per location check for normal distribution of square root transformed data
B6S02_sq <- my_data$square_root_transform[1:15]
B4S03_sq <- my_data$square_root_transform[16:30]
B4N01_sq <- my_data$square_root_transform[31:45]

# visual check
par(mfrow = c(2, 3))
qqPlot(B6S02)
qqPlot(B4S03)
qqPlot(B4N01)

qqPlot(B6S02_sq)
qqPlot(B4S03_sq)
qqPlot(B4N01_sq)

# check raw data
shapiro.test(B6S02)
shapiro.test(B4S03)
shapiro.test(B4N01)
# check square root transformation
shapiro.test(B6S02_sq)
shapiro.test(B4S03_sq)
shapiro.test(B4N01_sq)

#results: raw data for B4S03 is not normally distributed

```
# testing for homogeneity of variances
```{r}

leveneTest(my_data$square_root_transform, my_data$location, center = mean) 

# result: H0 (= no significant difference in variances between the locations) can't be rejected => variances are homogeneous => assumption is correct
```

# 1-way ANOVA
```{r}
anova = aov(square_root_transform~location, data=my_data)
summary(anova)

# result: null hypothesis can't be rejected, no significant differences?
```
# Tukey test
```{r}
test <- TukeyHSD(anova)
plot(test)
```


